### 2024-03-10 appendices/app_results_UQ.tex
Caste_study post-reviews by FC
Edit since UQ_TD_full has been completed

\Cref{tab:UQ_full} gives the ranking and total Sobol index over the total transition cost of each of the 34 parameters listed in \Cref{tab:UC_full}. The first column shows these indicators for the \gls{GSA} applied on the monthly pathway model that has some limitations \cite{limpens2023pathway} but has the main advantage to run much faster. The second column gives the same indicators but for an uncertainty quantification carried out on the hourly pathway model and only using the most impacting parameters. Given the similar rankings of the parameters between these two, this comparison shows that the monthly model can be a computationally efficient proxy to quantify the uncertainties of the actual hourly model and point out the key parameters on optimisation-driving objective, the total transition cost.

Besides the top-4 parameters, rankings are slightly different. However, this does not jeopardize the comparative analysis given the similar Sobol' indices. Given their wide range of uncertainty [-64.3\%; 179.8\%] and their significant role to meet the \ce{CO2}-budget, the cost of purchasing electrofuels is the first, by far, impacting parameter. Next, comes, naturally, the industrial \gls{EUD}, representing, at the nominal value, 60\% of the total demands by 2050. The top-3 is completed by the variation of the interest rate, directly impacting the annualisation and the salvage values of the assets. Finally, since the current Belgian whole-energy system deeply relies on fossil resources, and would still do in the near future, the cost of purchasing fossil fuels is part of the impacting parameters. On the contrary, due to the very low annualised, cost, long lifetime leading to a significant salvage value  and a low-emitting fuel, the parameters related to \gls{SMR} barely impact the total transition cost.

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost in the monthly and hourly pathway models. The similar rankings (and indices) show the validity of using the faster (even though less accurate) monthly model to assess uncertainties over the hourly model.}
\label{tab:UQ_full}
\centering
\begin{tabular}{l c c| c c}
\toprule
\multirow{2}{*}{\textbf{Parameter}}  & \multicolumn{4}{c}{\textbf{Ranking (Sobol' index)}}\\
 & \multicolumn{2}{c|}{Monthly model} 	& \multicolumn{2}{c}{Hourly model} \\ 	
\midrule
\textbf{Purchase electrofuels} & 1 & (47.4\%) & 1 & (44.4\%) \\
Industry EUD & 2 & (23.5\%) & 2 & (26.4\%)  \\
Interest rate & 3 & (11.0\%) &  3 & (13.2\%)  \\
Purchase fossil fuels  & 4 & (6.9\%) & 4 & (6.9\%)   \\
\midrule
Variable OPEX of technologies & 5 & (2.9\%) & 6 & (3.0\%) \\
Purchase biofuels & 6 & (2.6\%) & 5 & (3.0\%) \\
Hourly load factor PV & 7 & (1.9\%) & 9 & (1.3\%) \\
CAPEX electric motor & 8 & (1.9\%) & 7 & (2.5\%) \\
Hourly load factor wind turbines & 9 & (1.3\%) & 8 & (1.4\%) \\
Max capacity PV & 10 & (1.1\%) & 14 & (0.5\%) \\
\textbf{Potential capacity \gls{SMR}} & 11 & (0.9\%) & 11 & (1.0\%) \\
Available local biomass & 12 & (0.8\%) & 12 & (0.7\%) \\
CAPEX car & 13 & (0.7\%) & 10 & (1.2\%) \\
Passenger mobility EUD & 14 & (0.7\%) & 13 & (0.7\%) \\
\midrule
Modal share change LT-heat & 15 & (0.5\%) & \multicolumn{2}{c}{-} \\
Households EUD & 16 & (0.5\%) & \multicolumn{2}{c}{-} \\
Services EUD & 17 & (0.4\%) & \multicolumn{2}{c}{-} \\
Max capacity onshore wind & 18 & (0.3\%) & \multicolumn{2}{c}{-} \\
Max share of public transport & 19 & (0.3\%) & \multicolumn{2}{c}{-} \\
CAPEX PV & 20 & (0.2\%) & \multicolumn{2}{c}{-} \\
Max capacity offshore wind & 21 & (0.2\%) & \multicolumn{2}{c}{-} \\
Efficiency electric motor & 22 & (0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX fuel cell engine & 23 & (0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX ICE & 24 & (0.1\%) & \multicolumn{2}{c}{-} \\
Efficiency fuel cell engine & 25 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
Modal share change freight mobility & 26 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
Modal share change passenger mobility & 27 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX efficiency measures & 28 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX bus & 29 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX grid reinforcement & 30 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
CAPEX power grid & 31 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
\textbf{CAPEX \gls{SMR}} & 32 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
Available electricity import & 33 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
Purchase electricity & 34 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
\bottomrule

@@

% \begin{table}[htbp!]
% \caption{Total Sobol' indices of the uncertain parameters over the total transition cost in the monthly and hourly pathway models. The similar rankings (and indices) show the validity of using the faster (even though less accurate) monthly model to assess uncertainties over the hourly model.}
% \label{tab:UQ_full}
% \centering
% \begin{tabular}{l c c| c c}
% \toprule
% \multirow{2}{*}{\textbf{Parameter}}  & \multicolumn{4}{c}{\textbf{Ranking (Sobol' index)}}\\
%  & \multicolumn{2}{c|}{Hourly model} 	& \multicolumn{2}{c}{Model model} \\ 	
% \midrule
% Purchase electrofuels & 1 & (44.4\%) & 1 & (47.4\%) \\
% Industry EUD & 2 & (26.4\%) & 2 & (23.5\%)  \\
% Interest rate & 3 & (13.2\%) &  3 & (11.0\%)  \\
% Purchase fossil fuels & 4 & (6.9\%)  & 4 & (6.9\%)  \\
% \midrule
% Purchase biofuels & 5 & (3.0\%) & 6 & (2.6\%)  \\
% Variable OPEX of technologies & 6 & (3.0\%) & 5 & (2.9\%) \\
% CAPEX electric motor & 7 & (2.5\%) & 8 & (1.9\%)  \\
% Hourly load factor wind turbines & 8 & (1.4\%) & 9 & (1.3\%) \\ 
% Hourly load factor PV & 9 & (1.3\%) & 7 & (1.9\%)  \\
% CAPEX car & 10 & (1.2\%) & 13 & (0.7\%) \\
% \textbf{Potential capacity \gls{SMR}} & 11 & (1.0\%) & 11 & (0.9\%)  \\
% Available local biomass & 12 & (0.7\%)  & 12 & (0.8\%) \\
% Passenger mobility EUD & 13 & (0.7\%) & 14 & (0.7\%) \\
% Max capacity PV & 14 & (0.5\%) & 10 & (1.1\%)  \\
% \midrule
% Modal share change LT-heat & \multicolumn{2}{c}{-} & 15 & (0.5\%)  \\
% Households EUD & \multicolumn{2}{c}{-} & 16 & (0.5\%) \\
% Services EUD & \multicolumn{2}{c}{-} & 17 & (0.4\%) \\
% Max capacity onshore wind & \multicolumn{2}{c}{-} & 18 & (0.3\%) \\
% Max share of public transport & \multicolumn{2}{c}{-} & 19 & (0.3\%) \\
% CAPEX PV & \multicolumn{2}{c}{-} & 20 & (0.2\%) \\
% Max capacity offshore wind & \multicolumn{2}{c}{-} & 21 & (0.2\%)\\
% Efficiency electric motor & \multicolumn{2}{c}{-}  & 22 & (0.1\%) \\
% CAPEX fuel cell engine & \multicolumn{2}{c}{-}  & 23 & (0.1\%) \\
% CAPEX ICE & \multicolumn{2}{c}{-} & 24 & (0.1\%)\\
% Efficiency fuel cell engine & \multicolumn{2}{c}{-}  & 25 & ($<$0.1\%) \\
% Modal share change freight mobility & \multicolumn{2}{c}{-}  & 26 & ($<$0.1\%) \\
% Modal share change passenger mobility & \multicolumn{2}{c}{-}  & 27 & ($<$0.1\%)\\
% CAPEX efficiency measures & \multicolumn{2}{c}{-}  & 28 & ($<$0.1\%)\\
% CAPEX bus & \multicolumn{2}{c}{-}  & 29 & ($<$0.1\%)\\
% CAPEX grid reinforcement & \multicolumn{2}{c}{-} & 30 & ($<$0.1\%) \\
% CAPEX power grid & \multicolumn{2}{c}{-} & 31 & ($<$0.1\%)\\
% \textbf{CAPEX \gls{SMR}} & \multicolumn{2}{c}{-} & 32 & ($<$0.1\%) \\
% Available electricity import & \multicolumn{2}{c}{-}  & 33 & ($<$0.1\%) \\
% Purchase electricity  & \multicolumn{2}{c}{-} & 34 & ($<$0.1\%)\\
% \bottomrule

% \end{tabular}
% \end{table}

\newpage
\subsection{Imported renewable electrofuels}
\label{app:UQ_electrofuels}
\begin{table}[htbp!]
\caption{Comparison of the quantities of imported renewable electrofuels, in TWh, between the REF case, the SMR case and the statistical features from the \gls{GSA} (\ie Q1, median and Q3). 2020 is not in the table as, per assumption, no renewable electrofuel is imported for this year. For the sake of clarity, zeroes are replaced by ``-''.}
\label{tab:uq_ref_smr_med}
\begin{minipage}{\linewidth}
\centering
\begin{tabular}{l l | c c c c}
\toprule
\textbf{Year} & \textbf{Case} & \textbf{e-methane} & \textbf{e-hydrogen} & \textbf{e-ammonia} & \textbf{e-methanol}\\	
\toprule							
\multirow{5}{*}{2025}
 & REF & - & - & 10 & 52\\
 & SMR & - & -  & 10 & 29\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 2\\
 & Median & - & - & 10 & 46\\
 & Q3 & - & - & 11 & 55\\
\toprule
\multirow{5}{*}{2030}
 & REF & - & 1 & 10 & 52\\
 & SMR & - & 1 & 10 & 52\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 43\\
 & Median & - & - & 10 & 51\\
 & Q3 & - & 1 & 12 & 57\\
\toprule
\multirow{5}{*}{2035}
 & REF & - & 17 & 10 & 53\\
 & SMR & - & 17 & 10 & 53\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 44\\
 & Median & - & 4 & 11 & 52\\
 & Q3 & - & 16 & 32 & 58\\
\toprule
\multirow{5}{*}{2040}
 & REF & - & 16 & 23 & 54\\
 & SMR & - & 16 & 10 & 54\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 12 & 43\\
 & Median & - & 12 & 36 & 53\\
 & Q3 & 10 & 16 & 68 & 60\\
\toprule
\multirow{5}{*}{2045}
 & REF & 40 & 16 & 42 & 54\\
 & SMR & - & 16 & 11 & 54\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 24 & 43\\
 & Median & - & 13 & 49 & 52\\
 & Q3 & 44 & 17 & 77 & 60\\
\toprule
\multirow{5}{*}{2050}
 & REF & 39 & 16 & 44 & 55\\
 & SMR & 7 & 16 & 11 & 55\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 20 & 44\\
 & Median & 19 & 14 & 44 & 53\\
 & Q3 & 51 & 17 & 72 & 61\\
\bottomrule							
\end{tabular}
\end{minipage}
\end{table}

>>>>

\Cref{tab:UQ_full} gives the ranking and total Sobol index over the total transition cost of each of the 34 parameters listed in \Cref{tab:UC_full}. The first column shows these indicators for the \gls{GSA} applied on the hourly pathway model. For information, the second column gives the same indicators but for an uncertainty quantification carried out on the monthly pathway model that has some limitations \cite{limpens2024pathway} but has the main advantage to run much faster. Given the similar rankings of the parameters between these two, this comparison shows that the monthly model can be a computationally efficient proxy to quantify the uncertainties of the actual hourly model and point out the key parameters on optimisation-driving objective, the total transition cost.

Besides the top-4 parameters, rankings are slightly different. The main difference in terms of ranking relates to the import of electricity from abroad, \ie its cost of purchasing and its availability. Indeed, as observed by \citet{limpens2024pathway}, the monthly model does not require this import given the easier integration of monthly-averaged local \gls{VRES}. However, this does not jeopardize the comparative analysis given the similar Sobol' indices. Given their wide range of uncertainty [-64.3\%; 179.8\%] and their significant role to meet the \ce{CO2}-budget, the cost of purchasing electrofuels is the first, by far, impacting parameter. Next, comes, naturally, the industrial \gls{EUD}, representing, at the nominal value, 60\% of the total demands by 2050. The top-3 is completed by the variation of the interest rate, directly impacting the annualisation and the salvage values of the assets. Finally, since the current Belgian whole-energy system deeply relies on fossil resources, and would still do in the near future, the cost of purchasing fossil fuels is part of the impacting parameters. On the contrary, due to the very low annualised, cost, long lifetime leading to a significant salvage value and a low-emitting fuel, the parameters related to \gls{SMR} barely impact the total transition cost.

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost in the monthly and hourly pathway models. The similar rankings (and indices) show the validity of using the faster (even though less accurate) monthly model to assess uncertainties over the hourly model.}
\label{tab:UQ_full}
\centering
\begin{tabular}{l c c|c c}
\toprule
\multirow{2}{*}{\textbf{Parameter}}  & \multicolumn{4}{c}{\textbf{Ranking (Sobol' index)}}\\
& \multicolumn{2}{c|}{Hourly model} & \multicolumn{2}{c}{Monthly model}\\ 	
\midrule
\textbf{Purchase electrofuels} & 1 & (46.8\%) & 1 & (47.4\%) \\
Industry EUD & 2 & (23.2\%) & 2 & (23.5\%) \\
Interest rate & 3 & (12.0\%) & 3 & (11.0\%) \\
Purchase fossil fuels & 4 & (5.7\%) & 4 & (6.9\%) \\
\midrule
Variable OPEX of technologies & 5 & (3.1\%) & 5 & (2.9\%)\\
Purchase biofuels & 6 & (2.6\%) & 6 & (2.6\%) \\
CAPEX electric motor & 7 & (2.1\%) & 8 & (1.9\%) \\
Purchase electricity & 8 & (1.5\%) & 34 & ($<$0.1\%) \\
Hourly load factor wind turbines & 9 & (1.1\%) & 9 & (1.3\%)\\
Hourly load factor PV & 10 & (1.1\%) & 7 & (1.9\%) \\
\textbf{Potential capacity \gls{SMR}} & 11 & (0.9\%) & 11 & (0.9\%)\\
CAPEX car & 12 & (0.8\%) & 13 & (0.7\%) \\
Available local biomass & 13 & (0.7\%) & 12 & (0.8\%) \\
Passenger mobility EUD & 14 & (0.7\%) & 14 & (0.7\%) \\
Modal share change LT-heat & 15 & (0.5\%) & 15 & (0.5\%)\\
Max capacity PV & 16 & (0.5\%) & 10 & (1.1\%)\\
Households EUD & 17 & (0.5\%) & 16 & (0.5\%) \\
Services EUD & 18 & (0.5\%) &  17 & (0.4\%) \\
Max share of public transport & 19 & (0.3\%) & 19 & (0.3\%) \\
Max capacity onshore wind & 20 & (0.3\%) & 18 & (0.3\%) \\
CAPEX PV & 21 & (0.2\%) & 20 & (0.2\%) \\
Efficiency electric motor & 22 & (0.2\%) & 22 & (0.1\%)\\
Max capacity offshore wind & 23 & (0.2\%) & 21 & (0.2\%) \\
Available electricity import & 24 & (0.1\%) & 33 & ($<$0.1\%) \\
CAPEX ICE & 25 & (0.1\%) & 24 & (0.1\%)\\
CAPEX fuel cell engine & 26 & (0.1\%) & 23 & (0.1\%) \\
Efficiency fuel cell engine & 27 & (0.1\%) & 25 & ($<$0.1\%) \\
Modal share change freight mobility & 28 & (0.1\%) & 26 & ($<$0.1\%) \\
Modal share change passenger mobility & 29 & ($<$0.1\%) & 27 & ($<$0.1\%) \\
CAPEX grid reinforcement & 30 & ($<$0.1\%) & 30 & ($<$0.1\%) \\
CAPEX efficiency measures & 31 & ($<$0.1\%) & 28 & ($<$0.1\%) \\
CAPEX bus & 32 &($<$0.1\%)  & 29 & ($<$0.1\%) \\
\textbf{CAPEX \gls{SMR}} & 33 & ($<$0.1\%)  & 32 & ($<$0.1\%) \\
CAPEX power grid & 34 & ($<$0.1\%) & 31 & ($<$0.1\%) \\
\bottomrule

@@

%\begin{table}[htbp!]
%\caption{Total Sobol' indices of the uncertain parameters over the total transition cost in the monthly and hourly pathway models. The similar rankings (and indices) show the validity of using the faster (even though less accurate) monthly model to assess uncertainties over the hourly model.}
%\label{tab:UQ_full}
%\centering
%\begin{tabular}{l c c|c c| c c}
%\toprule
%\multirow{2}{*}{\textbf{Parameter}}  & \multicolumn{6}{c}{\textbf{Ranking (Sobol' index)}}\\
%& \multicolumn{2}{c}{Hourly model} & \multicolumn{2}{c|}{Monthly model} 	& \multicolumn{2}{c}{Hourly model} \\ 	
%\midrule
%\textbf{Purchase electrofuels} & 1 & (46.8\%) & 1 & (47.4\%) & 1 & (44.4\%) \\
%Industry EUD & 2 & (23.2\%) & 2 & (23.5\%) & 2 & (26.4\%)  \\
%Interest rate & 3 & (12.0\%) & 3 & (11.0\%) &  3 & (13.2\%)  \\
%Purchase fossil fuels & 4 & (5.7\%) & 4 & (6.9\%) & 4 & (6.9\%)   \\
%\midrule
%Variable OPEX of technologies & 5 & (3.1\%) & 5 & (2.9\%) & 6 & (3.0\%) \\
%Purchase biofuels & 6 & (2.6\%) & 6 & (2.6\%) & 5 & (3.0\%) \\
%CAPEX electric motor & 7 & (2.1\%) & 8 & (1.9\%) & 7 & (2.5\%) \\
%Purchase electricity & 8 & (1.5\%) & 34 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%Hourly load factor wind turbines & 9 & (1.1\%) & 9 & (1.3\%) & 8 & (1.4\%) \\
%Hourly load factor PV & 10 & (1.1\%) & 7 & (1.9\%) & 9 & (1.3\%) \\
%\textbf{Potential capacity \gls{SMR}} & 11 & (0.9\%) & 11 & (0.9\%) & 11 & (1.0\%) \\
%CAPEX car & 12 & (0.8\%) & 13 & (0.7\%) & 10 & (1.2\%) \\
%Available local biomass & 13 & (0.7\%) & 12 & (0.8\%) & 12 & (0.7\%) \\
%Passenger mobility EUD & 14 & (0.7\%) & 14 & (0.7\%) & 13 & (0.7\%) \\
%Modal share change LT-heat & 15 & (0.5\%) & 15 & (0.5\%) & \multicolumn{2}{c}{-} \\
%Max capacity PV & 16 & (0.5\%) & 10 & (1.1\%) & 14 & (0.5\%) \\
%Households EUD & 17 & (0.5\%) & 16 & (0.5\%) & \multicolumn{2}{c}{-} \\
%Services EUD & 18 & (0.5\%) &  17 & (0.4\%) & \multicolumn{2}{c}{-} \\
%Max share of public transport & 19 & (0.3\%) & 19 & (0.3\%) & \multicolumn{2}{c}{-} \\
%Max capacity onshore wind & 20 & (0.3\%) & 18 & (0.3\%) & \multicolumn{2}{c}{-} \\
%CAPEX PV & 21 & (0.2\%) & 20 & (0.2\%) & \multicolumn{2}{c}{-} \\
%Efficiency electric motor & 22 & (0.2\%) & 22 & (0.1\%) & \multicolumn{2}{c}{-} \\
%Max capacity offshore wind & 23 & (0.2\%) & 21 & (0.2\%) & \multicolumn{2}{c}{-} \\
%Available electricity import & 24 & (0.1\%) & 33 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX ICE & 25 & (0.1\%) & 24 & (0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX fuel cell engine & 26 & (0.1\%) & 23 & (0.1\%) & \multicolumn{2}{c}{-} \\
%Efficiency fuel cell engine & 27 & (0.1\%) & 25 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%Modal share change freight mobility & 28 & (0.1\%) & 26 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%Modal share change passenger mobility & 29 & ($<$0.1\%) & 27 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX grid reinforcement & 30 & ($<$0.1\%) & 30 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX efficiency measures & 31 & ($<$0.1\%) & 28 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX bus & 32 &($<$0.1\%)  & 29 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%\textbf{CAPEX \gls{SMR}} & 33 & ($<$0.1\%)  & 32 & ($<$0.1\%) & \multicolumn{2}{c}{-} \\
%CAPEX power grid & 34 & ($<$0.1\%) & 31 &  & \multicolumn{2}{c}{-} \\
%\bottomrule							
%
%\end{tabular}
%\end{table}

\newpage
\section{Imported renewable electrofuels}
\label{app:UQ_electrofuels}
\begin{table}[htbp!]
\caption{Comparison of the quantities of imported renewable electrofuels, in TWh, between the REF case, the SMR case and the statistical features from the \gls{GSA} (\ie Q1, median and Q3). 2020 is not in the table as, per assumption, no renewable electrofuel is imported for this year. For the sake of clarity, zeroes are replaced by ``-''.}
\label{tab:uq_ref_smr_med}
\begin{minipage}{\linewidth}
\centering
\begin{tabular}{l l | c c c c}
\toprule
\textbf{Year} & \textbf{Case} & \textbf{e-methane} & \textbf{e-hydrogen} & \textbf{e-ammonia} & \textbf{e-methanol}\\	
\toprule							
\multirow{5}{*}{2025}
 & REF & - & - & 10 & 52\\
 & SMR & - & -  & 10 & 29\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 2\\
 & Median & - & - & 10 & 47\\
 & Q3 & - & - & 11 & 55\\
\toprule
\multirow{5}{*}{2030}
 & REF & - & 1 & 10 & 52\\
 & SMR & - & 1 & 10 & 52\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 43\\
 & Median & - & - & 10 & 50\\
 & Q3 & - & 1 & 12 & 57\\
\toprule
\multirow{5}{*}{2035}
 & REF & - & 17 & 10 & 53\\
 & SMR & - & 17 & 10 & 53\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 9 & 42\\
 & Median & - & 5 & 11 & 51\\
 & Q3 & - & 16 & 33 & 58\\
\toprule
\multirow{5}{*}{2040}
 & REF & - & 16 & 23 & 54\\
 & SMR & - & 16 & 10 & 54\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 10 & 41\\
 & Median & - & 12 & 26 & 51\\
 & Q3 & 8 & 16 & 68 & 60\\
\toprule
\multirow{5}{*}{2045}
 & REF & 40 & 16 & 42 & 54\\
 & SMR & - & 16 & 11 & 54\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 12 & 42\\
 & Median & - & 12 & 37 & 52\\
 & Q3 & 35 & 17 & 71 & 60\\
\toprule
\multirow{5}{*}{2050}
 & REF & 39 & 16 & 44 & 55\\
 & SMR & 7 & 16 & 11 & 55\\
 \cmidrule{2 - 6}
 & Q1 & - & - & 11 & 43\\
 & Median & 1 & 13 & 31 & 53\\
 & Q3 & 36 & 17 & 66 & 61\\
\bottomrule							
\end{tabular}
\end{minipage}
\end{table}

### 2024-03-10 chap_RL/chap_RL.tex
Edit since UQ_TD_full has been completed

Defining the levers of action, the core of the policy, to support the transition of a country-size whole-energy system is challenging, especially when accounting for political and socio-technical aspects \cite{castrejon2020making}. In our work, focusing only on the techno-economic aspect, we assume that the actions taken by the agent are directly implemented and impacting the environment, without ``misfire''. In other words, considering only the techno-economic lens, there is no moderation nor contest towards the agent's actions, as the objective is to assess how far and when within the transition to push the different levers of action. Given the overall objective of the agent to succeed the transition, \ie respecting the \ce{CO2}-budget by 2050, we have defined the actions in this sense. The first action, $\mathrm{act}_{\mathrm{gwp}} \in [0,1]$, aims at limiting the emissions at the representative year ending the concerned time window, $\textbf{GWP\textsubscript{tot}}(y_{\text{end of the window}})$, between the level of emissions in 2020, \ie $\textbf{GWP\textsubscript{tot}}(2020)=123\,\text{Mt}_{\ce{CO2},\text{eq}}$, and carbon-neutrality:

@@

Properly defined the reward fed by the environment to the agent is crucial in \gls{RL} for several reasons.If the reward is not properly defined, the agent may optimize its policy for an unintended objective, leading to undesired or subotpimal behavior, \ie the so-called misalignment of the learning objective \cite{christiano2017deep}. Even worse, it can lead to reward hacking (or reward tampering) where the agent exploits loopholes in the reward function to achieve higher rewards without actually performing the desired task \cite{amodei2016concrete}. On the contrary, a proper definition of the reward function increases the sample efficiency, \ie requiring less episode to converge to the optimal policy.  It also makes the policy more stable and able to withstand variations and uncertainties in the environment \cite{henderson2018deep}.

@@

\newpage
\myparagraph{States}\\

Besides the reward, states are the other piece of information provided by the environment to the agent. In \gls{RL}, the purpose of states are to represent the current situation or configuration of the environment in which the agent operates. The primary function of states in RL is to provide the necessary context for the agent to choose appropriate actions based on its current observations and goals \cite{sutton2018reinforcement}. The challenge in the definition of the states is to provide enough information but not too much to avoid overwhelming the agent with non-informative features.

@@

\section{Results - As in Smart Energy Systems Conference}

>>>>

\noindent
Defining the levers of action, the core of the policy, to support the transition of a country-size whole-energy system is challenging, especially when accounting for political and socio-technical aspects \cite{castrejon2020making}. In our work, focusing only on the techno-economic aspect, we assume that the actions taken by the agent are directly implemented and impacting the environment, without ``misfire''. In other words, considering only the techno-economic lens, there is no moderation nor contest towards the agent's actions, as the objective is to assess how far and when within the transition to push the different levers of action. Given the overall objective of the agent to succeed the transition, \ie respecting the \ce{CO2}-budget by 2050, we have defined the actions in this sense. The first action, $\mathrm{act}_{\mathrm{gwp}} \in [0,1]$, aims at limiting the emissions at the representative year ending the concerned time window, $\textbf{GWP\textsubscript{tot}}(y_{\text{end of the window}})$, between the level of emissions in 2020, \ie $\textbf{GWP\textsubscript{tot}}(2020)=123\,\text{Mt}_{\ce{CO2},\text{eq}}$, and carbon-neutrality:

@@

\noindent
Properly defined the reward fed by the environment to the agent is crucial in \gls{RL} for several reasons.If the reward is not properly defined, the agent may optimize its policy for an unintended objective, leading to undesired or subotpimal behavior, \ie the so-called misalignment of the learning objective \cite{christiano2017deep}. Even worse, it can lead to reward hacking (or reward tampering) where the agent exploits loopholes in the reward function to achieve higher rewards without actually performing the desired task \cite{amodei2016concrete}. On the contrary, a proper definition of the reward function increases the sample efficiency, \ie requiring less episode to converge to the optimal policy.  It also makes the policy more stable and able to withstand variations and uncertainties in the environment \cite{henderson2018deep}.

@@

\myparagraph{States}\\

\noindent
Besides the reward, states are the other piece of information provided by the environment to the agent. In \gls{RL}, the purpose of states are to represent the current situation or configuration of the environment in which the agent operates. The primary function of states in RL is to provide the necessary context for the agent to choose appropriate actions based on its current observations and goals \cite{sutton2018reinforcement}. The challenge in the definition of the states is to provide enough information but not too much to avoid overwhelming the agent with non-informative features.

@@

\section{Learning process and convergence}
\label{sec:RL:learning}

Before testing the optimal policy $\pi^*\left(\bm{a}_n | \bm{o}_n\right)$, the first step consists in assessing the training of the \gls{NN}. For this, numerous episodes are played through the myopic optimization of the transition pathway of Belgium. At the beginning of each episode, the agent starts with the actual Belgian energy system of 2020 (see Appendix \ref{app:bel_2020}). Moreover, a sample of values is drawn for the uncertain parameters, affects the model for the 2020-2030 time window. This sample will remain valid for the following time windows. In other words, there is only one sample draw per episode (see Figure \ref{fig:ranges_transition}).  Then, the agent takes a set of actions, affecting the environment that feeds back the agent with a new state and a reward. This goes on until the end of the episode. Similarly to the \gls{UQ} analysis (see Section \ref{sec:meth:UQ}), the new sample of uncertain parameters for the new episode is drawn following the quasi-random Sobol' sampling technique \cite{bratley2003implementing}.

\begin{itemize}
\item Convergence of reward or success rate
\item Actions to take, which ones are binding
\item Show the state space: share of \gls{VRES} and efficiency
\item Assess the severity of learning on Monthly model, versus Hourly model. Compare the policies: one versus the other versus one followed by the other
\end{itemize}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.7\textwidth]{Binding_constr.pdf}
\caption{Binding versus non-binding constraints. In \gls{LP} where the feasibility domain is non-empty and bounded, the constraints defined a convex feasibility domain in the space of variables (here, x$_1$ and x$_2$). The optimal solution usually locates on a vertex of this domain, \ie the intersection of several constraints (here, constraints 2 and 3) limiting the solution. These constraints are considered as binding, \ie having a limiting impact on the optimal solution.}
\label{fig:Reward}
\end{figure}

\section{Testing and comparison with references}
\label{sec:RL:testing}
